---
title: "WGOA Survey Station-At Sea Outlook"
author: "Jesse F Lamb"
date: "`r Sys.Date()`"
output: html_document
output_dir: C:/Users/jesse.f.lamb/Work/GitHub/Survey_Station_Selection/R-files/2025 R Files/2025 Rendered Files 
---
Outlook: The purpose of this markdown is to determine an estimate of sampling time in order to adjust our plan to do the most sampling possible
with the DAS we have left.  Variables that can be adjusted:  Depart Time (line 86), Ship Speed = 10 (line 97).  Lastly, station order matters in terms of distance (line 60), so adjust if necessary.  

*Note: Be sure to properly enter the vector values of both "CompleteScenL" and "StaTDL", as they are needed to determine proper station number!
```{r Setup & Opening Table, message=FALSE, warning=FALSE, include=FALSE}

#Install Packages
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

packages <- c("tidyr",  "dplyr","lubridate","hms", "ggplot2", "ggmap", "lattice", "stringr", "data.table",  
              "tibble","readxl", "sf","maps", "mapdata", "mapplots", "mapview", "marmap", "Cairo", "ncdf4",
              "oce", "here","reshape2", "viridis", "export", "rnaturalearth", "kableExtra",
              "rnaturalearthdata", "forcats","sf", "geosphere")
ipak(packages)

# Create Base Layer Maps will all numbered stations labeled, with a more focused region (not whole WGOA)
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

```

## Overview:

Below are a Station Table and Map of the rest of the Survey

```{r Opening Table, echo=FALSE, warning=FALSE, message=FALSE}
# Load Station Data Table from DY19-05
#WGOAdata <- read.csv(here("Data","WGOA station_list_dougherty_2019_ProjInstrutions.csv"))
WGOAdata <- read.csv(here("Data","2025 Station Data","2025-04-01 WGOA DY25-05 GearTime & WPZ.csv"))
#WGOAdata$Station <- seq(0, nrow(WGOAdata) - 1)
#WGOAdata <- WGOAdata[,-1]

### Create Station & Runtime Dataset ############

###All Stations: based on "Station Column" of current WGOA Data
# in "Data"
AllSta<- c(0, 1:238) 

#Stations Left to Sample

# Completed Stations
CompleteScenL <- length(c(1:135)) #Fill in with completed StaTDL stations

# Staions Remaining 
StaTDL<-c(136:238)

### Cutting stations for time
# NE_Kodiak <- c(162:180)

#Drop Stations by subset using updated above vector, keep drops by date
#WGOAdata <- WGOAdata %>% subset(!(Station %in% STA.0516))

# Save CSV of above for Day to Day
#write.csv(WGOAdata, here("Data","DY23-07_Day2Day.csv"))

# Reverse Order Stations for Reference (DO NOT DELETE)
# 
#            c(1:9,24:28,59:53,84:77, 92:85, 100:93, 107:101,113:108,119:114, 
#           125:120, 131:126, 136:132,141:137, 142:142,152:148, 156:153, 159:157,165:160,
#           166:179,180,213:208,216:220,228:225,234:238,271)


ScenName <- "Stations Completed"
Scenario<-StaTDL # Always Change to Updated Stations Remaining
FinalSta<-data.frame("Station"=Scenario,"Stn_Order"= (CompleteScenL+1):(CompleteScenL+length(Scenario)))

# 1. Station Selection Through Scenario (above) 

Sta <- WGOAdata %>% left_join(.,FinalSta) %>%
  filter(Station %in% FinalSta$Station) %>%  
  arrange(Stn_Order)

# 2. Choose Departure Date
#DH_SIP <- ymd_hm("2023-05-17 14:00") 
#depart <- DH_SIP + dhours(9) # pre-survey calculations, takes ~8 hours to get to STA 1 from DH

# Use below during survey
depart <- lubridate::now(tzone = "America/Anchorage") + dhours(0.5) # Alaska time, for at sea calculations

# 3. Create data.frame that mirrors "Final Station Runtime.xlsx"=

SRTdata <- Sta %>% 
 mutate(Station_Time = lubridate::dhours(Total_Gear_Time)) %>% 
#<<<<<<< Updated upstream
  mutate(Ship_Speed = 11.5) %>%  # Ship speed is Variable
  mutate(lat_rad = LAT*pi/180) %>% mutate(long_rad = LON*pi/180) %>% 
  mutate(long_dif = long_rad - shift(long_rad)) %>% 
#>>>>>>> Stashed changes
  mutate(Dist_NM = ifelse(Station == 0,0,0) ) %>% 
  mutate(Dist_NM = distGeo(cbind(LON,LAT),cbind(shift(LON),shift(LAT)))/(1000*1.852)) %>% 
  # distGEO does what code below does in base r
#  mutate(Dist_NM = acos((sin(lag(lat_rad))*sin(lat_rad))+(cos(lag(lat_rad))*cos(lat_rad)*cos(long_dif)))/(pi/180)*60) %>%  
  mutate(Steam_T = Dist_NM/Ship_Speed) %>% 
  mutate(Steam_T = lubridate::dhours(Steam_T))

#Write .csv of new Station Order for Oscar Dyson crew 
ODsta <- SRTdata %>% select(Grid.ID,LAT,LON,Gear.Sampled,Stn_Order) 
#write.csv(ODsta, here("Data","For Dyson","DY23-07 Survey Station Order.csv"), row.names = FALSE) #Error if file open


# 4. rearrange data to calculate time to travel to and complete station operations
shipslog <- SRTdata %>% select(Station, Grid.ID, Station_Time, Steam_T, Stn_Order) %>% # SRTdata will be replaced by filtered stations above
  unite("Station",Station:Grid.ID, sep = "_", remove = FALSE) %>% 
  select(Station,Steam_T,Station_Time,Stn_Order) %>% 
  gather(key="Ops", Time_Dur, -Station,-Stn_Order) %>% arrange(Station) %>% 
  separate(Station, c("Station","Grid.ID"), sep = "_") %>% 
  transform(Station = as.numeric(Station)) %>% arrange(Stn_Order) 
shipslog <- na.omit(shipslog)

# 5. Create a table of Station Completion Times 
ShipsLog <- shipslog %>% group_by(Stn_Order, Station, Grid.ID) %>% 
  reframe(Time_Dur = lubridate::dseconds(sum(Time_Dur)))  %>% 
  mutate(Sta_Time =  depart) %>% mutate(cTime_Dur = cumsum(Time_Dur)) %>%
  mutate(Sta_Time = cTime_Dur + Sta_Time) %>% select(Stn_Order, Grid.ID,Sta_Time) %>% 
  rename("Station"="Stn_Order","Grid Station"="Grid.ID","Est.Station Time"="Sta_Time")

#6. Create a table to link Gear.Sampled for daily stations in ShipsLog to know Operations at each station
Sta_GearSamp <- SRTdata %>% select(Station,Grid.ID,LAT,LON,Gear.Sampled) %>% 
filter(Station %in% FinalSta$Station) %>% select(Grid.ID,LAT,LON,Gear.Sampled) %>% 
  rename("Grid Station"="Grid.ID")

ShipsLog <- left_join(ShipsLog,Sta_GearSamp, by = "Grid Station") %>% 
  select(Station,`Grid Station`,LAT,LON,Gear.Sampled,`Est.Station Time`)

# Output Station Table, include: Ships Log
kbl(ShipsLog, caption = "Next Stations", booktabs = T) %>% 
  kableExtra::kable_styling(bootstrap_options = "bordered", latex_options = c("striped", "HOLD_position"))
```

## Ship Track: Station Order


```{r Station Track Map,out.width = "100%", fig.align = "center", echo=FALSE, message=FALSE, warning=FALSE}

#SRTdata$LON <- SRTdata$LON*-1
# get bathymetry data

b = getNOAA.bathy(lon1 = min(SRTdata$LON)+-1, lon2 = max(SRTdata$LON)+1, 
                  lat1 = min(SRTdata$LAT)+-1, lat2 = max(SRTdata$LAT)+1, 
                 resolution = 15)
# convert bathymetry to data frame
bf = fortify.bathy(b)
# set map limits, whole region
lons = c(min(SRTdata$LON)+-1, max(SRTdata$LON)+1) #-140 = large map
lats = c(min(SRTdata$LAT)+-1,max(SRTdata$LAT)+1) # 52 = large map

# 5. Plot filtered station map with survey track
StaMapF <-ggplot()+geom_contour(data = bf, aes(x=x, y=y, z=z),breaks=c(-50),linewidth=c(0.5),colour="light grey")+ # add 50m contour
  geom_contour(data = bf, aes(x=x, y=y, z=z), breaks=c(-100), linewidth=c(0.5), colour="darkgrey")+ # add 100m contour
  geom_contour(data = bf, aes(x=x, y=y, z=z),breaks=c(-200),linewidth=c(0.3), colour="black")+ # add 200m contour
  geom_sf(data = world)+coord_sf(xlim = lons, ylim = lats, expand = FALSE)+ # add coastline
  #Plot points & station track
  geom_point(data = SRTdata, mapping = aes(LON, LAT, shape = Gear.Sampled), size = 1)+ 
  geom_path(data = SRTdata, mapping = aes(LON, LAT), linewidth = 0.25) +  
  geom_text(data = SRTdata, mapping = aes(LON, LAT, label = Stn_Order ), nudge_y = -0.08, size = 2.5)+ 
  scale_shape_discrete()+ theme_bw()+xlab("Longitude")+ylab("Latitude")+ ggtitle(paste(ScenName,length(Scenario),"Stations")) +
  theme(axis.text.x=element_text(size=8,hjust = 1,vjust = 1, angle = 45, color = "black"), axis.text.y = element_text(size=8, color = "black")) 

StaMapF
```


