---
title: "WGOA Survey Stations II"
author: "Jesse Lamb"
date: "2023-05-04"
output: html_document
---
## Overview:

This work is for planning of the DY23-07 Spring Larval Survey in the Western Gulf of Alaska (WGOA).  The survey has been allotted 12 days at sea (DAS), beginning from Dutch Harbor, AK and ending in Kodiak, AK.  This is the second iteration of this Rmd, in order to adjust station priority and  Gear.Sampled per station. 
```{r Setup & Opening Map, echo=TRUE, warning=FALSE, message=FALSE}
#Install Packages
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

packages <- c("tidyr",  "dplyr","lubridate","hms", "ggplot2", "ggmap", "lattice", "stringr", "data.table",  
              "tibble","readxl", "sf","maps", "mapdata", "mapplots", "mapview", "marmap", "Cairo", "ncdf4",
              "oce", "here","reshape2", "viridis", "export", "rnaturalearth", 
              "rnaturalearthdata", "forcats","sf", "geosphere")
ipak(packages) 

# Load Data: Using DY19-05 stations locations from that survey's Project Instructions, GS2 = Sparse Calvet grid
WGOAdata <- read.csv(here("Data","WGOA station_list_dougherty_2019_ProjInstrutions_GS2.csv"))
# If using "WGOA station_list_dougherty_2019_ProjInstrutions.csv" as WGOAdata, run the following 
# changing "Gear.Sampled" to reflect no neustons and Bongos at every station
WGOAdata <- WGOAdata %>% rename("LAT" = "Latitude.N.") %>% rename("LON"="Longitude..W.") %>% 
  mutate(Gear.Sampled = replace(Gear.Sampled,Gear.Sampled == "","BONGO")) %>% 
  mutate(Gear.Sampled = replace(Gear.Sampled,Gear.Sampled == "NEU","BONGO")) %>% 
  mutate(Gear.Sampled =  replace(Gear.Sampled,Gear.Sampled == "NEU, CTD","BONGO,CTD")) %>%
  mutate(Gear.Sampled = replace(Gear.Sampled,Gear.Sampled == "CTD","BONGO,CTD")) %>% 
  mutate(Gear.Sampled = replace(Gear.Sampled,Gear.Sampled == "CALVET","BONGO,CALVET")) %>% 
  mutate(Gear.Sampled = replace(Gear.Sampled,Gear.Sampled == "CALVET, CTD","BONGO,CALVET,CTD"))
WGOAdata$LON <- WGOAdata$LON*-1


# Create Base Layer Maps will all numbered stations labeled, with a more focused region (not whole WGOA)
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)
```

## R Markdown

Transforms "Station Run Time.xlsx" to calculate how long the survey could take based on Ship Speed = 11 knots and operations time estimates, which could be adjusted at sea.  
```{r station Spatial Limit, echo=TRUE, warning=FALSE, message=FALSE}
### Create Station & Runtime Dataset, mirroring "Final Station Runtime" Spreadsheets

AllSTA <- WGOAdata$Station 
ScenName <- "2023-05-16 Stations Possible"
Scenario<-AllSTA
FinalSta<-data.frame("Station"=Scenario,"Stn_Order"=1:length(Scenario))

# Station Selection Through Omission (Example Below)

# AllSta <- c(0:271) # All Possible Stations
# OmitSta <- c(0,60:225,271) # Stations to leave out
# FinalSta <- setdiff(AllSta,LLCvec) # All stations left
Sta <- WGOAdata %>% left_join(.,FinalSta) %>%
  filter(Station %in% FinalSta$Station) %>%  
  arrange(Stn_Order)


# 1. Create data.frame that mirrors "Final Station Runtime"
SRTdata <- Sta %>% 
  mutate(Bongo_hrs = ifelse(Gear.Sampled == "BONGO", 0.33,0)) %>% # BONGO at all stations
  mutate(BON_CTD_hrs = ifelse(Gear.Sampled == "BONGO,CTD",0.83,0)) %>% 
  mutate(BON_CalVET_hrs = ifelse(Gear.Sampled == "BONGO,CALVET", 0.66,0)) %>% 
  mutate(BON_Cal_CTD_hrs = ifelse(Gear.Sampled == "BONGO,CALVET,CTD",1.16,0)) %>% 
#<<<<<<< Updated upstream
  mutate(Station_Time = lubridate::dhours(rowSums(across(c(Bongo_hrs,BON_CTD_hrs,BON_CalVET_hrs,BON_Cal_CTD_hrs))))) %>% mutate(Ship_Speed = 11) %>% # Ship speed is Variable
  mutate(lat_rad = LAT*pi/180) %>% mutate(long_rad = LON*pi/180) %>% 
  mutate(long_dif = long_rad - shift(long_rad)) %>% 
#>>>>>>> Stashed changes
  mutate(Dist_NM = ifelse(Station == 0,0,0) ) %>% 
  mutate(Dist_NM = distGeo(cbind(LON,LAT),cbind(shift(LON),shift(LAT)))/(1000*1.852)) %>% # distGEO does what code below does in base
#  mutate(Dist_NM = acos((sin(lag(lat_rad))*sin(lat_rad))+(cos(lag(lat_rad))*cos(lat_rad)*cos(long_dif)))/(pi/180)*60) %>%  
  mutate(Steam_T = Dist_NM/Ship_Speed) %>% 
  mutate(Steam_T = lubridate::dhours(Steam_T))

# 2. rearrange data to calculate time to travel to and complete station operations
shipslog <- SRTdata %>% select(Station, Grid.ID,Station_Time, Steam_T, Stn_Order) %>% # SRTdata will be replaced by filtered stations above
  unite("Station",Station:Grid.ID, sep = "_", remove = FALSE) %>% 
  select(Station,Steam_T,Station_Time,Stn_Order) %>% 
  gather(key="Ops", Time_Dur, -Station,-Stn_Order) %>% arrange(Station) %>% 
  separate(Station, c("Station","Grid.ID"), sep = "_") %>% 
  transform(Station = as.numeric(Station)) %>% arrange(Stn_Order) 
shipslog <- na.omit(shipslog)
# 3. Choose Departure Date
depart <- ymd_hm("2023-05-14 12:00")
# 4. Create a table of Station Completion Times 
ShipsLog <- shipslog %>% group_by(Stn_Order, Station, Grid.ID) %>% 
  reframe(Time_Dur = lubridate::dseconds(sum(Time_Dur)))  %>% 
  mutate(Sta_Time =  depart) %>% mutate(cTime_Dur = cumsum(Time_Dur)) %>%
  mutate(Sta_Time = cTime_Dur + Sta_Time) %>% select(Station, Stn_Order, Grid.ID,Sta_Time) %>% 
  left_join(.,Grid.loc) %>% select(Stn_Order,Grid.ID,LAT,LON,Sta_Time)


```

## MAIN STATION GRID EDITing FOR "Outlook" and "Day to Day" Pre Cruise

Create Station Table file from "SRTdata" above, with a column listing station sampling priority: "A" = high, "B" = moderate, "C" = low.  This will allow easy filtering at sea if/when we lose time for expedient decision making.  
```{r Station Table,echo=TRUE, warning=FALSE, message=FALSE }

# Create table, remove all CTDs not part of transects; create "A,B,C" station vectors
# Note: Station 0 = Dutch Harbor, 271 = Kodiak Pier

# Below is an EXAMPLE to get code to run and plot proper (4/27/23)
A <- c(0,24:28, 53:59, 77:84,93:100,108:113,120:125,132:141,148:152,160:165,
       167,169,171,173,175,177,208:213,225:228,220,218,216,234:238,271) 
B <- c(1:9, 38:44,68:76,85:92,101:107,114:119,126:131,142:147,153:159,166,168,170,
       172,174,176,178,179,180)
Port <-c(0)
# C = anything not A & B, see line 121 below
#moved rom B to C <- c(181, 201:197, 224:221,219,217,215,214,229:233,239,243:245,
#       249:251,255:257,261:263,267:269)


WGOA_StaTable <- WGOAdata %>% select(Station, Grid.ID,LAT,LON,Gear.Sampled) %>% 
  mutate(Gear.Sampled = ifelse(Station == 24,"BONGO",Gear.Sampled)) %>%
  mutate(Gear.Sampled = ifelse(Station == 40,"BONGO",Gear.Sampled)) %>% 
  mutate(Gear.Sampled = ifelse(Station == 94,"BONGO",Gear.Sampled)) %>% 
  mutate(Gear.Sampled = ifelse(Station %in% Port,"",Gear.Sampled)) %>% 
  mutate(Sta_Pri = if_else(Station %in% A,"A",if_else(Station %in% B,"B","C"))) 

## Remove C priority Stations; filter dropped stations from Outlook 
# Outlook Stations left, in order
StaTDL<-c(1:9,24:28,59:53,84:77, 92:85, 100:93, 107:101,113:108,119:114, 
          125:120,131:126,136:132,141:137,142:147,152:148, 156:153, 159:157,165:160,
          166:179,180,213:208,216,218,220,228:225,234:238,271)

WGOA_StaTable <- WGOA_StaTable %>% filter(Sta_Pri != "C") %>% 
  filter(Station %in% StaTDL)
  
ScenName <- "DY23-07 All A & B Priority Stations"

write.csv(WGOA_StaTable, here("Data","DY23-07_A&B_priority.csv"))

```

## Station Priority Map


```{r Station Track Map,out.width = "100%", fig.align = "center", echo=FALSE, message=FALSE, warning=FALSE}

#WGOA_StaTable$LON <- WGOA_StaTable$LON*-1
# get bathymetry data

#CVgrid <- c(114:142)
#WGOA_StaTable <- WGOA_StaTable %>% filter(Station %in% CVgrid)
#ScenName <- "CalVetGrid Stations"

b = getNOAA.bathy(lon1 = min(WGOA_StaTable$LON)+-1, lon2 = max(WGOA_StaTable$LON)+1, 
                  lat1 = min(WGOA_StaTable$LAT)+-1, lat2 = max(WGOA_StaTable$LAT)+1, 
                 resolution = 15)
# convert bathymetry to data frame
bf = fortify.bathy(b)
# set map limits, whole region
lons = c(min(WGOA_StaTable$LON)+-1, max(WGOA_StaTable$LON)+1) #-140 = large map
lats = c(min(WGOA_StaTable$LAT)+-1,max(WGOA_StaTable$LAT)+1) # 52 = large map

# Create Color Vector: stoplight
stpliteCol <- c("blue","orange","red")

# Plot filtered station map with survey track
StaMapF <-ggplot()+geom_contour(data = bf, aes(x=x, y=y, z=z),breaks=c(-50),linewidth=c(0.5),colour="light grey")+ # add 50m contour
  geom_contour(data = bf, aes(x=x, y=y, z=z), breaks=c(-100), linewidth=c(0.5), colour="darkgrey")+ # add 100m contour
  geom_contour(data = bf, aes(x=x, y=y, z=z),breaks=c(-200),linewidth=c(0.3), colour="black")+ # add 200m contour
  geom_sf(data = world)+coord_sf(xlim = lons, ylim = lats, expand = FALSE)+ # add coastline
  #Plot points & station track
  geom_point(data = WGOA_StaTable, mapping = aes(LON, LAT, shape = Gear.Sampled, color= Sta_Pri), size = 2)+ 
  #geom_path(data = WGOA_StaTable, mapping = aes(LON, LAT), linewidth = 0.25) +  
  geom_text(data = WGOA_StaTable, mapping = aes(LON, LAT, label = Station ), nudge_y = -0.08, size = 2.5)+ 
  scale_shape_discrete()+ theme_bw()+xlab("Longitude")+ylab("Latitude")+
  scale_color_manual(values = stpliteCol)+
    ggtitle(paste(ScenName)) +
  theme(axis.text.x=element_text(size=8,hjust = 1,vjust = 1, angle = 45, color = "black"), axis.text.y = element_text(size=8, color = "black")) 

StaMapF

ggsave(paste0(ScenName,".png"),path = here("Docs"), height = 8.5, width = 11, units = "in")
```
