---
title: "WGOA Survey Stations"
author: "Jesse F Lamb"
date: "`r Sys.Date()`"
output: html_document
---

## Overview:

This work is for planning of the DY23-07 Spring Larval Survey in the Western Gulf of Alaska (WGOA).  The survey has been allotted 12 days at sea (DAS), beginning from Dutch Harbor, AK and ending in Kodiak, AK.  The first maps will focus on the most relevant spatial regions, due to prior years sampling.  The subsequent maps will delve into the most efficient cruise tracks give our limited DAS compared to previous years (21 DAS).  
```{r Setup & Opening Map, echo=TRUE, warning=FALSE, message=FALSE}
#Install Packages
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

packages <- c("tidyr",  "dplyr","lubridate","hms", "ggplot2", "ggmap", "lattice", "stringr", "data.table",  
              "tibble","readxl", "sf","maps", "mapdata", "mapplots", "mapview", "marmap", "Cairo", "ncdf4",
              "oce", "here","reshape2", "viridis", "export", "rnaturalearth", 
              "rnaturalearthdata", "forcats","sf", "geosphere")
ipak(packages) 

# Create a base Layer map of the core sampled areas, as per JL and LR meeting on 2023-02-16
# Create Base Layer Maps will all numbered stations labeled, with a more focused region (not whole WGOA)
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

# Read in "late-larval core" region used for larval time-series, as shapefile:
# convert to Lat/lon projection.
shp.mp <- read_sf(dsn=here("Data","mapping_data","NewLateLarvalPolygon"),layer="Survey_poly")
#shp.mp.LL<-st_transform(shp.mp,CRS("+proj=longlat"))
shp.mp.LL<-st_transform(shp.mp,crs=("+proj=longlat"))

WGOAdata <- read.csv(here("Data","WGOA station_list_dougherty_2019_ProjInstrutions.csv"))

# Count Gear.Sampled
table(WGOAdata$Gear.Sampled)

WGOAdata$Longitude..W. <- WGOAdata$Longitude..W.*-1
# get bathymetry data
b = getNOAA.bathy(lon1 = -166, lon2 = -140, lat1 = 52, lat2 = 62, 
                  resolution = 15)
## Querying NOAA database ...
## This may take seconds to minutes, depending on grid size
## Building bathy matrix ...
# convert bathymetry to data frame
bf = fortify.bathy(b)
# get regional polygons
reg = map_data("world2Hires")
reg = subset(reg, region %in% c('USSR', 'USA'))
# convert lat longs
reg$long = (360 - reg$long)*-1
# set map limits, whole region
lons = c(-167, -149) #-140 = large map
lats = c(53, 60.25) # 52 = large map

######################################
# make plot
GOARegion_Map_New <- ggplot()+
   # add 50m contour
  geom_contour(data = bf, 
               aes(x=x, y=y, z=z),
               breaks=c(-50),
               linewidth=c(0.5),
               colour="light grey")+
  # add 100m contour
  geom_contour(data = bf, 
               aes(x=x, y=y, z=z),
               breaks=c(-100),
               linewidth=c(0.5),
               colour="darkgrey")+
  # add 200m contour
  geom_contour(data = bf, 
               aes(x=x, y=y, z=z),
               breaks=c(-200),
               linewidth=c(0.5),
               colour="black")+
  
  # add late larval "core" polygon
 # geom_sf(data = shp.mp.LL, color="green", fill=NA, linewidth = 1.5) +
  
  # add coastline
  geom_sf(data = world)+
  coord_sf(xlim = lons, ylim = lats, expand = FALSE)+
  
  
  #Plot WGOA points
  geom_point(data = WGOAdata, mapping = aes(Longitude..W., Latitude.N., shape = Gear.Sampled), size = 1)+ 
  #Delete "shape..." for just stations
  geom_text(data = WGOAdata, mapping = aes(Longitude..W., Latitude.N., label = Station ), 
            nudge_y = -0.08, size = 2.5)+ 
  #station text size & nudge NOT IN 'aes' configure projection and plot domain coord_map(xlim = lons, ylim = lats)+
  
  # formatting
  scale_shape_discrete()+ 
  theme_bw()+
  xlab("Longitude")+
  ylab("Latitude")+
  theme(axis.text.x=element_text(size=10, color = "black"), axis.text.y = element_text(size=10, color = "black")) 
  
 GOARegion_Map_New

ggsave("SpringLarval_AllSTA_2023_Gear&StationNo.png",path = here("Docs"), height = 8.5, width = 11, units = "in")

```

The plot above represents the sampling plan developed from the DY19-05 survey, which had 21 DAS.  We will narrow down this amount of stations to represent the core area (**define here**), Shelikof Strait, and the NE portion of Kodiak Is.  We will have multiple iterations of plans, where we will plot potential survey tracks for our given time of 12 DAS and the starting port of Dutch, Harbor AK.    


```{r station Spatial Limit, echo=TRUE, warning=FALSE, message=FALSE}
### Create Station & Runtime Dataset, mirroring "Final Station Runtime" Spreadsheets

# Filter out Stations 
# LLC = "late larval core" from polygon plot above
LLCvec <- c(0,60:225,271)
Scen1<-c(0,24:28,44:38,53:59,60:179, 180, 181,201:197,208:213, 216,218,220,228:225,
         234:238,239, 245:243,249:251,257:255,261:263, 269:267,271)

Scenario<-Scen1
FinalSta<-data.frame("Station"=Scenario,"Stn_Order"=1:length(Scenario))

# Station Selection Through Omission (Example Below)

# AllSta <- c(0:271) # All Possible Stations
# OmitSta <- c(0,60:225,271) # Stations to leave out
# FinalSta <- setdiff(AllSta,LLCvec) # All stations left
Sta <- WGOAdata %>% left_join(.,FinalSta) %>%
  filter(Station %in% FinalSta$Station) %>%
  arrange(Stn_Order)

# 1. Create data.frame that mirrors "Final Station Runtime"
SRTdata <- Sta %>% rename(LAT = Latitude.N., LON = Longitude..W.) %>% 
  mutate(Gear.Sampled = replace(Gear.Sampled,Gear.Sampled == "NEU","")) %>% 
  mutate(Gear.Sampled =  replace(Gear.Sampled,Gear.Sampled == "NEU, CTD","CTD")) %>%  
  mutate(Bongo_hrs = ifelse(Station > 0, 0.33,0)) %>% 
  mutate(CTD_hrs = ifelse(Gear.Sampled == "CTD",0.5,0)) %>% 
  mutate(CalVET_hrs = ifelse(Gear.Sampled == "CALVET", 0.33,0)) %>% 
  mutate(CalCTD_hrs = ifelse(Gear.Sampled == "CALVET, CTD",0.83,0)) %>% 
#<<<<<<< Updated upstream
  mutate(Station_Time = lubridate::dhours(rowSums(across(c(Bongo_hrs,CTD_hrs,CalVET_hrs,CalCTD_hrs))))) %>% mutate(Ship_Speed = 11) %>% # Variable
  mutate(lat_rad = LAT*pi/180) %>% mutate(long_rad = LON*pi/180) %>% 
  mutate(long_dif = long_rad - shift(long_rad)) %>% 
#>>>>>>> Stashed changes
  mutate(Dist_NM = ifelse(Station == 0,0,0) ) %>% 
  mutate(Dist_NM = distGeo(cbind(LON,LAT),cbind(shift(LON),shift(LAT)))/(1000*1.852)) %>% # distGEO does what code below does in base
#  mutate(Dist_NM = acos((sin(lag(lat_rad))*sin(lat_rad))+(cos(lag(lat_rad))*cos(lat_rad)*cos(long_dif)))/(pi/180)*60) %>%  
  mutate(Steam_T = Dist_NM/Ship_Speed) %>% 
  mutate(Steam_T = lubridate::dhours(Steam_T))

# 2. rearrange data to calculate time to travel to and complete station operations
shipslog <- SRTdata %>% select(Station, Grid.ID, Station_Time, Steam_T, Stn_Order) %>% # SRTdata will be replaced by filtered stations above
  unite("Station",Station:Grid.ID, sep = "_", remove = FALSE) %>% 
  select(Station,Steam_T,Station_Time,Stn_Order) %>% 
  gather(key="Ops", Time_Dur, -Station,-Stn_Order) %>% arrange(Station) %>% 
  separate(Station, c("Station","Grid.ID"), sep = "_") %>% 
  transform(Station = as.numeric(Station)) %>% arrange(Stn_Order) 
shipslog <- na.omit(shipslog)
# 3. Choose Departure Date
depart <- ymd_hm("2023-05-14 12:00")
# 4. Create a table of Station Completion Times 
ShipsLog <- shipslog %>% group_by(Stn_Order, Station, Grid.ID) %>% 
  reframe(Time_Dur = lubridate::dseconds(sum(Time_Dur)))  %>% 
  mutate(Sta_Time =  depart) %>% mutate(cTime_Dur = cumsum(Time_Dur)) %>%
  mutate(Sta_Time = cTime_Dur + Sta_Time) %>% select(Station, Stn_Order, Grid.ID,Sta_Time)

# 5. Plot filtered station map with survey track
StaMapF <-ggplot()+geom_contour(data = bf, aes(x=x, y=y, z=z),breaks=c(-50),linewidth=c(0.5),colour="light grey")+ # add 50m contour
  geom_contour(data = bf, aes(x=x, y=y, z=z), breaks=c(-100), linewidth=c(0.5), colour="darkgrey")+ # add 100m contour
  geom_contour(data = bf, aes(x=x, y=y, z=z),breaks=c(-200),linewidth=c(0.3), colour="black")+ # add 200m contour
  geom_sf(data = world)+coord_sf(xlim = lons, ylim = lats, expand = FALSE)+ # add coastline
  #Plot points & station track
  geom_point(data = SRTdata, mapping = aes(LON, LAT, shape = Gear.Sampled), size = 1)+ 
  geom_path(data = SRTdata, mapping = aes(LON, LAT), linewidth = 0.25) +  
  geom_text(data = SRTdata, mapping = aes(LON, LAT, label = Station ), nudge_y = -0.08, size = 2.5)+ 
  scale_shape_discrete()+ theme_bw()+xlab("Longitude")+ylab("Latitude")+ ggtitle("Late Larval Core to 225") +
  theme(axis.text.x=element_text(size=10, color = "black"), axis.text.y = element_text(size=10, color = "black")) 

StaMapF

ggsave("SpringLarval_ST60_ST225_2023_Gear&StationNo.png",path = here("Docs"), height = 8.5, width = 11, units = "in")


```

## Potential Station spatial and Temporal Layouts

Below we will show potential station maps (station plots, w/ lines from station to station), along with time tables for each station map.  

```{r pressure, echo=FALSE}
### Function Below for use to compare multiple maps at the same time. 

### Create Station Mapping function.  This will use filtered station data from the entire dataset (WGOA Data).  

# StaMapF <- function(z,zz), where z= station position data & zz = name of filtered data, (i.e. EOS = "every other station")
# StaMapF mirrors same format as station and gear maps above w/ lines linking stations 
StaMapF <- function(zed,zz){ggplot()+geom_contour(data = bf, aes(x=x, y=y, z=z),breaks=c(-50),linewidth=c(0.5),colour="light grey")+ # add 50m contour
  geom_contour(data = bf, aes(x=x, y=y, z=z), breaks=c(-100), linewidth=c(0.5), colour="darkgrey")+ # add 100m contour
  geom_contour(data = bf, aes(x=x, y=y, z=z),breaks=c(-200),linewidth=c(0.5), colour="black")+ # add 200m contour
  geom_sf(data = world)+coord_sf(xlim = lons, ylim = lats, expand = FALSE)+ # add coastline
  #Plot Z points
  geom_point(data = z, mapping = aes(LON, LAT, shape = Gear.Sampled), size = 1)+ 
  geom_path(data = z, mapping = aes(LON, LAT,)) +  
  geom_text(data = z, mapping = aes(LON, LAT, label = Station ), nudge_y = -0.08, size = 2.5)+ 
  scale_shape_discrete()+ theme_bw()+xlab("Longitude")+ylab("Latitude")+ 
  theme(axis.text.x=element_text(size=10, color = "black"), axis.text.y = element_text(size=10, color = "black")) }
                                          
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
